---
title: "Data Wrangling with dplyr and tidyr"
author :
- Aboubacar HEMA 

date: "June 26, 2020"
output:
  word_document:
    reference_docx: Reference.docx
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Import with readr, tibble, and tidyr

R's **tidyverse** is built around **tidy data** stored in
**tibbles**, an enhanced version of a data frame.

## Read functions 

### Read tabular data to tibbles

These functions share the common arguments:
**read_**
```{r message=FALSE, warning=FALSE}

```

#### read_csv()

Reads comma delimited files. 


# Tidy Data
Tidy Data is  a foundation for wrangling in R. In a tidy data set, each variable is saved in its own column and each observation is saved in its own row.Tidy data complements R's vectorized
operations. R will automatically preserve
observations as you manipulate variables.
No other format works as intuitively with R.

## Syntax - Helpful conventions for wrangling
### tbl_df()

Converts data to tbl class. tbl's are easier to examine than
data frames. R displays only the data that fits on screen.
```{r echo=TRUE, message=FALSE, warning=FALSE}
dplyr::tbl_df(iris)
```

### glimpse()

Information dense summary of tbl data.
```{r echo=TRUE, message=FALSE, warning=FALSE}
dplyr::glimpse(iris)
```

### View()

View data set in spreadsheet-like display (note capital V).

```{r echo=TRUE, message=FALSE, warning=FALSE}
#utils::View(iris)
```

### %>%

Passes object on lef hand side as first argument (or .
argument) of function on righthand side. 
x %>% f(y) is the same as f(x, y) and y %>% f(x, ., z) is the same as f(x, y, z ).
"Piping" with %>% makes code more readable, e.g:
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
iris %>%
 group_by(Species) %>%
 summarise(avg = mean(Sepal.Width)) %>%
 arrange(avg)
```

## Reshaping Data - Change the layout of a data set

### gather()

Gather columns into rows.
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyr)
mini_iris <-
  iris %>%
  group_by(Species) %>%
  slice(1)
mini_gather<-mini_iris %>% gather(key = "flower_att", value = "measurement", -Species)
#utils::View(mini_iris)
```

### spread()

Spread rows into columns
```{r echo=TRUE, message=FALSE, warning=FALSE}
mini_gather %>% spread(flower_att,measurement)
```

### separate()

Separate one column into several.

```{r message=FALSE, warning=FALSE}
library(dplyr)
df <- data.frame(x = c(NA, "a.b", "a.d", "b.c"))
df
#utils::View(df)
df %>% separate(x, c("A", "B"))
```

If you just want the second variable:
```{r}
df %>% separate(x, c(NA, "B"))
```

### unite()

Unite several columns into one.
```{r}
df<-df %>% separate(x, c("A", "B"))
df %>% unite("x", A:B, remove = FALSE)
```

To remove missing values:
```{r}
df %>% unite("x", A:B, na.rm = TRUE, remove = FALSE)

```

## Subset Observations (Rows)

### filter()

Extract rows that meet logical criteria.

### distinct()
Remove duplicate rows. 

### sample_frac()

Randomly select fraction of rows.

### sample_n()

Randomly select n rows. 

### slice()

Select rows by position. 

### top_n()

Select and order top n entries (by group if grouped data).



## Subset Variables (Columns)

### select()

Select columns by name or helper function.Helper functions for select.

#### select(iris,contains("."))

Select columns whose name contains a character string. 

#### select(iris, ends_with("Length"))

Select columns whose name ends with a character string. 

 
#### select(iris, everything())

Select every column.

#### select(iris, matches(".t."))

Select columns whose name matches a regular expression.

#### select(iris, num_range("x", 1:5))

Select columns named x1, x2, x3, x4, x5.

#### select(iris, one_of(c("Species", "Genus")))

Select columns whose names are in a group of names.

#### select(iris, starts_with("Sepal"))

Select columns whose name starts with a character string.

#### select(iris, Sepal.Length:Petal.Width)

Select all columns between Sepal.Length and Petal.Width (inclusive).

#### select(iris, -Species)

Select all columns except Species. 

## Summarise Data

### summarise()

Summarise data into single row of values. 

### summarise_each(iris,funs(mean))

Apply summary function to each column.

### count()

Count number of rows with each unique value of
variable (with or without weights).

## Make New Variables

### mutate()

Compute and append one or more new columns.

### mutate_each(iris,funs(min_rank))

Apply window function to each column. 

### transmute()

Compute one or more new columns. Drop original columns.

## Combine Data Sets

### Mutating Joins

#### left_join()

Join matching rows from b to a.

#### right_join()

Join matching rows from a to b.

#### inner_join()

Join data. Retain only rows in both sets.

#### full_join()

Join data. Retain all values, all rows.

### Filtering Joins

#### semi_join()

All rows in **a** that have a match in **b**.

#### anti_join()

All rows in **a** that do not have a match in **b**.

### Set Operations

#### intersect()

Rows that appear in both **y** and **z**.

#### union()

Rows that appear in either or  both **y** and **z**.

#### setdiff()
Rows that appear in  **y** but not **z**.

### Binding

#### bind_rows()

Append **z** to **y** as new rows.

#### bind_cols()

Append z to y as new columns.
Caution: matches rows by position.

## Group Data

### group_by(iris,Species)

Group data into rows with the same value of Species. 

### ungroup(iris)

Remove grouping information from data frame. 

### iris %>% group_by(Species) %>% summarise(.)

Compute separate summary row for each group.

### iris %>% group_by(Species) %>% mutate(.)

Compute new variables by group.
